---
title: "H"
author: "Gio Jacuzzi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Acoustic Entropy (H)

#### Sueur et al 2008

TODO

#### Bradfer-Lawrence et al 2019

Soundscape patterns (H) - Increases with greater evenness of amplitude among frequency bands and/or time steps. Returns a value between 1 (an even signal, either noisy across frequency bands or completely silent) and 0 (a pure tone with all energy in one frequency band).

Patterns in this study (H) - Highest values come from near-silent recordings, with no wind, and only faint bird calls. Lowest values produced when insect noise dominated a single frequency band.

Principle (Alcocer et al 2022):

-   Ht (Temporal Entropy Index) - Shannon entropy of the amplitude envelope
-   Hf (Spectral Entropy Index) - Shannon entropy of the frequency spectrum
-   H (Acoustic Entropy Index) - Product of Ht and Hf

```{r}
library(soundecology)
library(tuneR)
library(seewave)

# Load test data
path = '~/Desktop/oesf-examples'
files = list.files(path=path, pattern='*.wav', full.names=T, recursive=T)

results = data.frame()
for (file in files) {
  wav = readWave(file)
  h = H(wav)
  results = append(results, h)
}
```

### References
