---
title: "NDSI"
author: "Gio Jacuzzi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Normalized Difference Soundscape Index (NDSI)

#### Kasten et al 2012

The goal of our normalized difference soundscape index (NDSI) is to estimate the level of anthropogenic disturbance on the soundscape by computing the ratio of human-generated (anthrophony) to biological (biophony) acoustic components found in field collected sound samples. As shown in Fig. 5, the analysis of a large number of recordings collected at several locations, revealed that mechanical sounds are most prevalent between 1 and 2 kHz and biological sounds are most prevalent between 2and8kHz(Gage and Napoletano, 2004; Gage et al., 2001).

To compute the overall level of biophony present in an acoustic signal, we first compute the power spectral density (PSD) (Welch,1967) of the signal. Then, a rectangular estimate of the PSD integral is computed for the anthropogenic and biophonic frequency ranges, and NDSI computed as

$$
NDSI=\frac{\beta - \alpha}{\beta + \alpha}
$$

where $\beta$ and $\alpha$ are the total estimated PSD for the largest 1 kHz biophony bin and the anthrophony bin respectively. The NDSI for the soundscape at a location does not remain constant, and changes according to time of day or day of year can be used to plot how NDSI changes over time. Note that NDSI is a ratio in the range [−1 to + 1], where + 1 indicates a signal containing no anthrophony. However, for some biophonic vocalizations a low NDSI can also indicate the presence of certain types of animals. For instance, the common loon (Gavia immer) has a low frequency call that often produces a NDSI score less than −0.8. Such anomalies are expected when attempting to characterize the highly variable nature of soundscape acoustics, and indicates that advancements are needed to help characterize and search acoustic observations. However, searching by NDSI can be a useful filter to help limit the number of recordings for further examination.

#### Wildlife Acoustics

-   Quantifies disturbance based on the ratio of biotic to anthropogenic sound in recordings
-   Also uses area under the curve, but divides the power spectrum into two parts (anthrophony, typically 1-2kHz vs biophony, often 2-8kHz).
-   Kasten et al had a big library of acoustic data and used it to develop the NDSI
-   It's important to understand that there are limitations
-   NDSI = (biophony - anthrophony) / biophony + anthrophony)
-   Tending to -1 = more anthrophony
-   Tending to +1 = more biophony

#### Bradfer-Lawrence et al 2019

Soundscape patterns - The ratio of acoustic energy between two defined frequency bands. Designed to estimate the level of anthropogenic disturbance in the soundscape. This index gives the ratio between anthrophony and biophony. This relies on the theoretical frequency split between the two facets of the soundscape. In some habitats, anthrophony is restricted to the 1 -- 2 kHz range, while most biophony appears in the 2 -- 8 kHz range. However, NDSI is sensitive to low frequency biophony. NDSI values range from -1 to +1, with +1 indicating no anthrophony in the soundscape. Following (Fuller et al. 2015), the upper biophony range was extended from 8 kHz to 11 kHz, to incorporate higher frequency sounds typical of tropical habitats (Sueur et al. 2014).\*

Patterns in this study - High values reflect high levels of insect biophony, with minimal noise in the 1-2 kHz range. Low values arise when insect biophony dominates the 1-2 kHz band.

Principle: Ratio of the spectral density of two frequency bands (i.e. anthrophony and biophony) (Alcocer et al 2022)

\*NDSI can be calculated using the seewave function "NDSI". Outputs are not identical from the seewave and soundecology packages, probably because of the way in which the amplitude values are calculated. Moreover, "NDSI" (seewave) may have incorrectly implemented the original formulation as it uses 2000 -- 8999 Hz instead of 2 -- 8 kHz. In addition, when analysing stereo recordings, seewave only returns a single averaged value across channels rather than a separate value for each channel as in soundecology. In both "ndsi" and "NDSI" the amplitude values for biophony are summed across all 1 kHz frequency bands, whereas in Kasten et al. (2012) the index is described as using the largest 1kHz biophony band. Summing the biophony bands reduces apparent differences among recordings, so that only those with particularly dominant anthrophony have low NDSI values.

```{r}
library(soundecology)
library(tuneR)
library(seewave)
source('helpers.R')

# Load test data
path = '~/Desktop/oesf-examples/short'
files = list.files(path=path, pattern='*.wav', full.names=T, recursive=T)

# Parameters
f1_min = 0
f1_max = 2000
f2_min = 2000
f2_max = 8000
window = 512  # FFT window size

results = data.frame()

for (file in files) {
  message(file)
  sound = readWave(file)
  if (dc_offset(sound@left)) sound@left = dc_correct(sound@left)
  
  # Calculate index
  ndsi_soundecology = ndsi(
    soundfile = sound,
    fft_w     = window,
    anthro_min = f1_min,
    anthro_max = f1_max,
    bio_min    = f2_min,
    bio_max    = f2_max
  )
  ndsi_seewave = NDSI(soundscapespec( # TODO: investigate discrepancies
      wave = sound,
      wl   = window,
      plot = FALSE
    ),
    anthropophony = (f1_min/1000):(f1_max/1000),
    biophony      = (f2_min/1000):(f2_max/1000)
  )
  results = rbind(results, data.frame(
    soundecology=ndsi_soundecology$ndsi_left, seewave=ndsi_seewave))
  
  
  # Plot
  p = spectrogram(sound, alim=c(-55,0), flim=c(0,12), tlim=c(0,10)) +
    labs(subtitle = paste0('Frequency band limits [', f1_min, ',', f1_max, ') and [', f2_min, ',', f2_max, '] Hz')) +
    annotate('rect', xmin = 0, xmax = 10, ymin = f1_min/1000, ymax = f1_max/1000-.05,
           alpha = .2, color = 'red', fill = 'red') +
    annotate('rect', xmin = 0, xmax = 10, ymin = f2_min/1000, ymax = f2_max/1000,
           alpha = .1, color = 'white', fill = 'white') +
    theme(legend.position='none', text = element_text(size = 30))
  print(p + scale_y_continuous(limits=c(0,12), n.breaks=7))
  ggsave(p + scale_y_continuous(limits=c(0,12), n.breaks=7), file='~/Desktop/ndsi.png', width=12, height=6)
}
```

```{r}
# Benchmarking soundecology against seewave
library(microbenchmark)

benchmark = microbenchmark(
  ndsi( # soundecology
    soundfile = sound,
    fft_w     = window,
    anthro_min = f1_min,
    anthro_max = f1_max,
    bio_min    = f2_min,
    bio_max    = f2_max
  ),
  NDSI(soundscapespec( # seewave
      wave = sound,
      wl   = window,
      plot = FALSE
    ),
    anthropophony = c(f1_min/1000, f1_max/1000),
    biophony      = c(f2_min/1000, f2_max/1000)
  )
)
benchmark
autoplot(benchmark)
```

### References
