---
title: "AR"
author: "Gio Jacuzzi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Acoustic Richness Index (AR)

#### Depraetere et al 2021

Principle: Weighted product of the ranks of M and H$_{t}$ indices (Alcocer et al 2022)

The acoustic entropy $H$ index was successfully tested in a tropical forest with high signal-to-noise ratio. However, in the context of a temperate habitat, the acoustic activity may be lower and background noise from wind and rain higher, potentially biasing $H_{f}$, generating false high values when wind and rain dominate over animal calls.

Acoustic Richness $AR$ elaborates on $H_{t}$ and $M$ following

$$
AR=\frac{\text{rank}(H_{t})*\text{rank}(M)}{n^{2}}, \text{ with } 0 \leq AR \leq 1
$$

Note that computing AR on a set of a single object does not make any sense.

```{r}
library(tuneR)
library(seewave)
source('helpers.R')

# Load test data
path = '~/Desktop/oesf-examples/short'
files = list.files(path=path, pattern='*.wav', full.names=T, recursive=T)

results = data.frame()
for (file in files) {
  sound = readWave(file)
  # TODO
}
```

### References
